---
title: "ResumeMatch Lab — Parallel Statistical Validation (R)"
format:
  html:
    toc: true
    embed-resources: true
execute:
  warning: false
  message: false
---

## Purpose

This notebook reproduces ResumeMatch Lab's **entire statistical pipeline in R**, on the
*identical* paired-delta vector exported from the Python engine
(`analysis/r/demo_deltas.csv`, written by `scripts/export_demo_deltas.py`). If Python and R
agree on the same data, the statistics are implementation-independent. The tolerance table
lives in `docs/python-vs-r.md`.

## Setup

```{r}
#| label: setup
library(readr)
library(dplyr)
library(boot)
library(pwr)

d <- read_csv("demo_deltas.csv", show_col_types = FALSE)
delta <- d$delta
n <- length(delta)
n
```

## 1. Effect size & frequentist tests (normality-gated)

```{r}
mean_delta <- mean(delta)
cohens_d <- mean_delta / sd(delta)

# Mirror the Shapiro-Wilk gate (R caps the test at n = 5000).
samp <- if (n > 5000) sample(delta, 5000) else delta
shapiro_p <- shapiro.test(samp)$p.value

t_res <- t.test(delta, mu = 0)                       # paired t == one-sample on delta
w_res <- wilcox.test(delta, mu = 0, exact = FALSE, correct = FALSE)

list(mean_delta = mean_delta, cohens_d = cohens_d, shapiro_p = shapiro_p,
     t_p = t_res$p.value, wilcox_p = w_res$p.value)
```

## 2. Bootstrap CI — 10,000 resamples, percentile + BCa

```{r}
set.seed(42)
bs <- boot(delta, statistic = function(x, i) mean(x[i]), R = 10000)
boot.ci(bs, type = c("perc", "bca"))
```

## 3. Power analysis

```{r}
achieved_power <- pwr.t.test(n = n, d = abs(cohens_d), sig.level = 0.05,
                             type = "one.sample")$power
required_n_80 <- pwr.t.test(d = abs(cohens_d), sig.level = 0.05, power = 0.8,
                            type = "one.sample")$n
c(achieved_power = achieved_power, required_n_80 = required_n_80)
```

## 4. CUPED — lm() residualization on job-side covariates

```{r}
m <- lm(delta ~ factor(cluster_id) + scale(desc_len), data = d)
variance_reduction <- summary(m)$r.squared            # == 1 - var(resid)/var(delta)
variance_reduction
```

## 5. mSPRT — Robbins mixture, always-valid p-value

```{r}
s2 <- var(delta); tau2 <- s2
ns <- seq_len(n); xbar <- cumsum(delta) / ns
lambda <- sqrt(s2 / (s2 + ns * tau2)) *
          exp((ns^2 * tau2 * xbar^2) / (2 * s2 * (s2 + ns * tau2)))
always_valid_p <- cummin(pmin(1, 1 / lambda))
tail(always_valid_p, 1)
```

## 6. Bayesian Beta-Binomial posterior

```{r}
k <- sum(delta > 0)
post_a <- 1 + k; post_b <- 1 + n - k
list(k = k,
     posterior_mean = post_a / (post_a + post_b),
     credible_interval = qbeta(c(0.025, 0.975), post_a, post_b),
     prob_b_beats_a = 1 - pbeta(0.5, post_a, post_b))
```

## 7. Per-cluster tests + multiple-comparison correction

```{r}
per <- d %>%
  group_by(cluster_id, cluster_label) %>%
  summarise(n = n(), mean_delta = mean(delta),
            p_raw = t.test(delta, mu = 0)$p.value, .groups = "drop")
per$p_bonferroni <- p.adjust(per$p_raw, method = "bonferroni")
per$p_bh         <- p.adjust(per$p_raw, method = "BH")
per
```

## Reproducibility

Identical input vector to Python, `set.seed(42)` for the bootstrap. Render with
`quarto render analysis/r/compare.qmd`. Compare the printed values against the Python
reference in `docs/python-vs-r.md`.
